---
title: "Course Project Reproductible Analysis"
author: "Flahault Grégoire"
date: "11/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##I. Loading and preprocessing the data
We start by opening the data file and turning the date column into date data.
```{r, echo=TRUE}
setwd("/Users/gregoireflahault/Desktop/Coursera/Rep_Data_W2")
dat<-read.csv(file = "activity.csv")
dat$date<-as.Date(dat$date, "%Y-%m-%d")
par(mfrow=c(1,1))
```

##II. What is mean total number of steps taken per day?
We sum the number of steps for each day into the agg variable.
Then we plot the frequency of the number of steps
```{r, echo=TRUE}
agg<-aggregate(dat$steps, list(date= dat$date), FUN=sum)
hist(agg$x, breaks = 10, main= "histogram of the total number of steps taken each day", 
                         xlab = "number of steps")
```

Then i calculate the mean and the median values of those sums
```{r, echo=TRUE}
mean(agg$x, na.rm=TRUE)
median(agg$x, na.rm=TRUE)
```

##III. What is the average daily activity pattern?
Then we calculate the mean number of step for each interval over the 2 months
And we plot this.
```{r, echo=TRUE}
agg_2<-aggregate(dat$steps, list(interval = dat$interval), FUN=mean, na.rm=TRUE)
with( agg_2, plot(x = interval, y = x, 
             type = "l",
             main = "average number of steps taken for each interval",
             ylab = "number of steps"))
```

The interval with the highest mean number of steps
```{r, echo=TRUE}
agg_2$interval[agg_2$x==max(agg_2$x)]
```

##IV. Imputing missing values
The number of missing values into our data.set (number of missing rows)
```{r, echo=TRUE}
sum(is.na(dat$steps))
```

Now i create a new data.set called dat_2
To this new data.set, i substitute all the missing values by the mean value of steps
for the particular interval.
This allows us to calculate a new total of steps for each day that might be more relevant.

```{r, echo=TRUE}
dat_2<-dat
dat_2$steps_mean<-agg_2$x

dat_2$full<-dat_2$steps
dat_2$full[is.na(dat_2$steps)==TRUE]<-dat_2$steps_mean[is.na(dat_2$steps)==TRUE]

agg_3<-aggregate(dat_2$full, list(date= dat_2$date), FUN=sum)
```

We can now plot the new frequency for each number of steps per day
```{r, echo=TRUE}
hist(agg_3$x, breaks = 10, main = "histogram of the total number of steps taken each day",
                           xlab = "numbers of steps")
```

The new calculation of the mean give us:
```{r, echo=TRUE}
mean(agg_3$x)
median(agg_3$x)
```
Which is not far away from what we got previously

##V. Are there differences in activity patterns between weekdays and weekends?
I add a new column to study if each day is a *week-day* or a *week-end*
```{r, echo=TRUE}
#download.packages("timeDate", "/Users/gregoireflahault/Desktop/Coursera/Rep_Data_W2")
#install.packages("timeDate")
library(timeDate)
dat_2$day<-isWeekday(dat_2$date)
dat_2$dday<-"week-end"
dat_2$dday[dat_2$day==TRUE]<-"week-day"
```

By subsetting in 2 different groups we can plot 2 graphs depending on the day 
of the week.
Then for each group we aggregate to have the mean number of steps for each interval.
```{r, echo=TRUE, fig.height=10}
wday<-subset(dat_2, dat_2$dday=="week-day")
      j<-aggregate(wday$full, list(interval= wday$interval), FUN= mean)
wend<-subset(dat_2, dat_2$dday=="week-end")
      k<-aggregate(wend$full, list(interval= wend$interval), FUN= mean)

par(mfrow= c(2,1))
with(j, plot(x = interval, y = x , type="l",
             main = "week-day mean number of steps per interval",
             ylab = "number of steps"))
with(k, plot(x = interval, y = x , type="l",
             main = "week-end mean number of steps per interval",
             ylab = "number of steps"))

```
